---
title: "Spicy Like Curry - To shoot or not to shoot?"
author: "Nils Indreiten"
format: 
  html:
    theme: pulse
    code-fold: true
    code-summary: "Show the code"
    toc: true
---

## Executive Summary

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(BasketballAnalyzeR,tidyverse,zoo,gt,hoopR,patchwork,tidymodels,reactablefmtr,htmltools,crosstalk,patchwork,shiny,vip)

team_logos <- readr::read_csv("Logo_info.csv") %>% 
  select(TEAM_LOGO,TEAM_NAME,CONFERENCE) %>% 
  unique() %>% 
 rbind(c("https://upload.wikimedia.org/wikipedia/en/thumb/2/21/Portland_Trail_Blazers_logo.svg/2560px-Portland_Trail_Blazers_logo.svg.png","Portland Trail Blazers", "Western")) %>% 
  mutate(TEAM_NAME=if_else(TEAM_NAME == "Los Angeles Clippers","LA Clippers",TEAM_NAME)) 

final_res <- readRDS("~/Documents/R/spicy_like_curry/final_res.rds")
final_xgb <- readRDS("~/Documents/R/spicy_like_curry/final_xgb.rds")
xgb_results <- readRDS("~/Documents/R/spicy_like_curry/xgb_results.rds")

nba_team_box <- read_csv("nba_team_box.csv")
nba_player_box <- read_csv("nba_player_box.csv")
nba_pbp <- read_csv("nba_pbp.csv")
```

The idea for this project came from a conversation with a friend, we were discussing
whether it is possible to predict whether a basketball player will make a shot or not. 
In specific, the question was whether _Stephen Curry_ would make a shot.
This project aims to predict whether Stephen Curry will make a shot 
at a play-by-play level, based on a variety of features. 

The project is structured as follows: First general trends in the NBA
are explored, honing in on the point guard position, which is Curry's
position. Then a few key features are selected and an XGBoost classifier
is trained and tested. The model performance is then evaluated in unseen
data, season 2024. The model was trained and tested on data from the 2022 
and 2023 seasons. Finally, the findings are concluded and future research 
directions are suggested.

## Exploratory Data Analysis & Data Wrangling

### Data Collection

Since Stephen Curry is a point guard, let's focus on the trends for that 
position. First, let's take a look at the overall stats for all point 
guards in the NBA. The _hoopR_ package let's users retrieve NBA data at different
levels of granularity, for this we need data at all levels of aggregations, therefore
we retrieve team box, player box, and play by play data:

```{r}
#| eval: false

# NBA team box data, 2022-2024 seasons:
nba_team_box <- hoopR::load_nba_team_box(2022:hoopR::most_recent_nba_season())

# NBA player box data, 2022-2024 seasons:
nba_player_box <- hoopR::load_nba_player_box(2022:hoopR::most_recent_nba_season())

# NBA play by play data, 2022-2024 seasons:
nba_pbp <- hoopR::load_nba_pbp(2022:hoopR::most_recent_nba_season())
```

We now have three dataframes, each denoting different information, then `nba_team_box`
dataframe is the most broad, followed by `nba_player_box` and `nba_pbp` which are more 
granular. Next, using the `nba_player_box` dataframe as the starting point, we filter
only for point guards and then aggregate the data to get a sense of the overall performance
by player and team:

```{r}
#| message: false

# Team names data frame:

team_names <- nba_team_box %>% 
  select(team_name,team_display_name) %>% 
  unique()

# Filter nba_player_box, for point guards

nba_player_box %>% 
  filter(athlete_position_name == "Point Guard") -> point_guard_df

# Select columns that we need 
point_guard_df %>% 
  select(season,athlete_id,athlete_display_name,athlete_position_name,team_name,
         minutes,field_goals_made,three_point_field_goals_made,free_throws_made,
         rebounds,assists,offensive_rebounds,defensive_rebounds,steals,blocks) -> pg_stats

# PG stats summary  

pg_stats %>% 
  filter(season == 2023 & !grepl("Team",team_name)) %>%
  group_by(athlete_display_name, team_name) %>% 
  mutate(total_points =    sum(
    field_goals_made + three_point_field_goals_made + free_throws_made,
    na.rm = TRUE
  )) %>% 
  ungroup() %>%
  group_by(athlete_display_name, team_name) %>% 
  summarise(
  minutes = sum(minutes, na.rm = TRUE),
  field_goals_made = mean(field_goals_made, na.rm = TRUE),
  three_point_field_goals_made = mean(three_point_field_goals_made, na.rm = TRUE),
  free_throws_made = mean(free_throws_made, na.rm = TRUE),
  assists = mean(assists, na.rm = TRUE),
  rebounds = mean(offensive_rebounds + defensive_rebounds, na.rm = TRUE),
  total_points = sum(max(total_points))) %>%
  arrange(desc(total_points)) %>% 
  ungroup() -> pg_stats

pg_stats %>% 
  head(5) 
```

The resulting data frame contains a summary of the player performance for the 
2023 season, summarised by player & team. Let's wrap the dataframe in a reactive
framework, so that you (the reader), can filter and slice the data as you wish, 
where `AFG` refers to average field goals made, `3PAFG` refers to average three 
point made, `AFT` refers to average free throws made, `AST` refers to average assists,
`REB` refers to average rebounds, and `TPTS` refers to total points:

```{r}
#| message: false
#| warning: false
#| echo: false

# Reactive set up
pg_stats %>% 
  left_join(team_names, by = c("team_name" = "team_name")) %>% 
  left_join(team_logos, by = c("team_display_name" = "TEAM_NAME")) %>% 
  select(athlete_display_name,TEAM_LOGO,everything()) ->pg_stats

### create shared dataset for crosstalk
crosstalk_data <- SharedData$new(pg_stats)

### crosstalk team filter
team_filter <- filter_select(
  id = "team",
  label = "TEAM",
  sharedData = crosstalk_data,
  group = ~ team_display_name
)

### crosstalk conference filter
conference_filter <- filter_select(
  id = "conf",
  label = "CONFERENCE",
  sharedData = crosstalk_data,
  group = ~ CONFERENCE
)

### crosstalk minutes filter
minutes_filter <- filter_slider(
  id = "minutes",
  label = "MINUTES PLAYED",
  sharedData = crosstalk_data,
  column = ~ minutes,
  ticks = TRUE,
  dragRange = FALSE,
  step = 500,
  width = "60%"
)

### load font from google fonts
htmltools::tags$link(href = "https://fonts.googleapis.com/css?family=Chivo:400,600,700&display=swap", rel = "stylesheet")

nba_table <- reactable(
  crosstalk_data,
  theme = fivethirtyeight(centered = TRUE),
  compact = TRUE,
  columnGroups = list(
    colGroup(name = "OVERALL STATS", columns = c("field_goals_made","three_point_field_goals_made","free_throws_made","assists","rebounds"))
  ),
  showSortIcon = FALSE,
  searchable = TRUE,
  language = reactableLang(
    searchPlaceholder = "SEARCH FOR A PLAYER..."
  ),
  defaultPageSize = 10,
  columns = list(
    team_display_name = colDef(show = FALSE),
    team_name = colDef(show = FALSE),
    CONFERENCE = colDef(show = FALSE),
    TEAM_LOGO = colDef(
      name = "Team",
      maxWidth = 70,
      align = "center",
      cell = embed_img(height = 25, width = 40),
      style = list(borderRight = "1px solid #777")
    ),
    athlete_display_name = colDef(minWidth = 150,
                                  name = "Player",
                                  # merge the "Name" column with the "Postion" column and place it below
                                  cell = merge_column(pg_stats, "team_name", merged_position = "below"),
                                  style = list(borderRight = "1px solid #777")),
    minutes = colDef(
      name = "Minutes",
      maxWidth = 85,
      align = "center",
      cell = icon_assign(
        pg_stats,
        icon = "stopwatch",
        fill_color = "#555555",
        buckets = 5
      ),
      style = list(borderRight = "1px dashed rgba(0, 0, 0, 0.3)")
    ),
    field_goals_made = colDef(
      name = "AFG",
      maxWidth = 60,
      style = color_scales(pg_stats, colors = c("#fd84a9", "white", "#42c2ca")),
      format = colFormat(digits =2)
    ),
    three_point_field_goals_made = colDef(
      name = "3PAFG",
      maxWidth = 60,
      style = color_scales(pg_stats, colors = c("#fd84a9", "white", "#42c2ca")),
      format = colFormat(digits =2)
    ),
    free_throws_made = colDef(
      name = "AFT",
      maxWidth = 60,
      style = color_scales(pg_stats, colors = c("#fd84a9", "white", "#42c2ca")),
      format = colFormat(digits =2)
    ),
    assists = colDef(
      name = "AST",
      maxWidth = 60,
      style = color_scales(pg_stats, colors = c("#fd84a9", "white", "#42c2ca")),
      format = colFormat(digits =2)
    ),
    rebounds = colDef(
      name = "REB",
      maxWidth = 60,
      style = color_scales(pg_stats, colors = c("#fd84a9", "white", "#42c2ca")),
      format = colFormat(digits =2)
    ),
    total_points = colDef(
      name = "TPTS",
      minWidth = 150,
      align = "center",
      cell = data_bars(
        pg_stats,
        fill_color = c("#de425b", "#ec9c9d", "#8cbcac", "#488f31"),
        number_fmt = scales::comma_format(),
        round_edges = TRUE,
        border_style = "solid",
        border_color = "gold",
        border_width = ".8px",
        text_position = "above"),
      style = list(borderLeft = "1px dashed rgba(0, 0, 0, 0.3)")
    )
  )
)

### display crosstalk filters
div(
  style = "display: flex;",
  div(style = "margin-right: 10px; width: 80%;", team_filter),
  div(style = "margin-right: 10px; width: 60%;", conference_filter),
  div(style = "width: 150%;", minutes_filter)
)

### display table
div(nba_table)
```

In terms of total points it looks like Stephen Curry comes in at 4th place, whilst
Trae Young comes out on top. However, if we focus on the `3PAFG` metric, Curry 
comes out on top. Let's hone in on the top 6 point guards in terms of 2-pointers, 
3-pointers, assists, rebounds and free throws made, scaling to accommodate the 
different magnitudes of each metric:

```{r}
pg_stats %>% 
  select(team_name,athlete_display_name,field_goals_made,three_point_field_goals_made,free_throws_made,rebounds,assists) %>%
  rename(
    Team = team_name,
    Player = athlete_display_name,
    P2M = field_goals_made,
    P3M = three_point_field_goals_made,
    FTM = free_throws_made,
    REB = rebounds,
    AST = assists) %>% 
  mutate(Player = case_when(Team == "Nets" & Player == "Kyrie Irving" ~ "Kyrie Irving (BKN)",
                            Team == "Mavericks" & Player == "Kyrie Irving" ~ "Kyrie Irving (DAL)",
                            TRUE ~ Player)) %>%
  as.data.frame() %>% 
  arrange(desc(P2M)) %>% 
  slice_head(n = 6) -> Pbox.PG
  
attach(Pbox.PG)
X <- data.frame(P2M, P3M, FTM, REB, AST)

detach(Pbox.PG)
radialprofile(data=X, title=Pbox.PG$Player, std=TRUE,ncol.arrange = 3)
```

The first thing to note is that there are two entries for Kyrie Irving, one for 
the Brooklyn Nets and one for the Dallas Mavericks. This is due to a transfer 
sometime in the season, either way Irving's performance in the Nets is comparable to his 
performance in Dallas, perhaps slightly better in Dallas at an overall level. Curry clearly 
differentiates himself in terms of 3-pointers, relative to the other point guards. 
Luka Doncic is more of a rebound and assist heavy point-guard, similarly Damian 
Lillard overindexes in assists. Shai Gilgeous Alexander stands out in terms of 
free throws made.  Lets look at the shot chart for Stephen Curry in the 2023 season
excluding free throws, how many shots did he make and miss?

```{r}
nba_pbp %>%  
  filter(season == 2023) %>% 
  select(coordinate_x_raw,coordinate_y_raw,scoring_play,clock_minutes) %>% 
  mutate(result=as.factor(if_else(scoring_play == TRUE, "made","missed")))-> curry_shots


subdata <- curry_shots %>% as.data.frame()
subdata$xx <- subdata$coordinate_x_raw-25
subdata$yy <- subdata$coordinate_y_raw-44

shotchart(data=subdata, x="xx", y="yy",z="result" ,type=NULL,scatter=TRUE)+
  theme(legend.position = "bottom")
```

We see a concentration of field goals made near the board, as well as quite an evenly
distribution of shots along the 3-point line. There also seems to be a higher concentration
of misses along the 3-point line, suggesting that whilst Curry might be making the most three points 
he is also missing a high number of 3-point shots. Let's look at the density of shots made using a 
heatmap:

```{r}
#| warning: false
shotchart(data=subdata, x="xx", y="yy", type="density-raster",scatter=FALSE)
```

This reveals a similar pattern, with a high density of shots made near the board
and a large number of shots made along the top of the 3-point line. 

## Modelling

For this modelling exercise we will consider the play-by-play data and try to 
predict the outcome of a play, excluding free throw plays for Stephen Curry. 
we will need to do some data wrangling to retrieve some key information at the 
play-by-play level, such as the opponent athlete name, game id and filtering 
for all seasons equal to or less than 2023. These steps are outlined below:

First we get a distinct list of players and game IDs for the Golden State 
Warriors (Steph Curry's team), this will help up us match the opponent athlete
and the respective play-by-play data:

```{r}
# Get the distinct set of player ids:
nba_player_box %>%
  select(athlete_id,athlete_display_name) %>% 
  distinct() %>%
  arrange(athlete_id)-> player_ids

player_ids %>% 
  group_by(athlete_id) %>%
  mutate(row = row_number()) %>%
  filter(row==1) %>% 
  ungroup() %>% 
  select(athlete_id,athlete_display_name) -> player_ids

# Get Game IDs for the Golden State Warriors
nba_team_box %>% 
  filter(season <= 2023 & team_display_name == "Golden State Warriors") %>%
  select(game_id,opponent_team_name) -> game_ids
```

Second, we  combine the `player_ids` and `game_ids` dataframes with the play-by-play
data, giving us the base data for the modelling:

```{r}
#| eval: FALSE
nba_pbp %>%  
  filter(season <= 2023, 
           athlete_id_1 == 3975 & # Filter for Curry
           shooting_play == TRUE & # And only shooting plays
           !grepl("Free Throw",type_text)) %>% # exclude free throws
  select(game_id,coordinate_x_raw,coordinate_y_raw,scoring_play,
         clock_minutes, athlete_id_2,clock_minutes,clock_seconds,qtr) %>% 
  left_join(player_ids,by=c("athlete_id_2"="athlete_id")) %>% 
  inner_join(game_ids, by=join_by(game_id)) %>% 
  select(-athlete_id_2) %>% 
  rename(oponent_athlete=athlete_display_name) %>% 
  mutate(oponent_athlete = if_else(is.na(oponent_athlete),"no direct oponent",oponent_athlete),
         scoring_play = fct_rev(as_factor(scoring_play))) -> base_steph_pbp
```


```{r}
#| echo: false

nba_pbp %>%  
  filter(season <= 2023 ) %>%  
          # athlete_id_1 == 3975 & # remove this filter
          # shooting_play == TRUE & # and this one
          # !grepl("Free Throw",type_text)) %>% # exclude free throws
  select(game_id,coordinate_x_raw,coordinate_y_raw,scoring_play,
         clock_minutes, athlete_id_2,clock_minutes,clock_seconds,qtr) %>% 
  left_join(player_ids,by=c("athlete_id_2"="athlete_id")) %>% 
  inner_join(game_ids, by=join_by(game_id)) %>% 
  select(-athlete_id_2) %>% 
  rename(oponent_athlete=athlete_display_name) %>% 
  mutate(oponent_athlete = if_else(is.na(oponent_athlete),"no direct oponent",oponent_athlete),
         scoring_play = fct_rev(as_factor(scoring_play))) -> base_steph_pbp
base_steph_pbp
```

With our base dataframe, we can now create the training, testing and cross-validation
folds:

```{r}
set.seed(123)
steph_split <- initial_split(base_steph_pbp, prop = 0.8, strata = scoring_play)
steph_train <- training(steph_split)
steph_test <- testing(steph_split)

steph_folds <- vfold_cv(steph_train, v = 10, strata = scoring_play)
```

With the base data in place, we are now able to create the splits for training and testing
the model. We will use a 80/20 split for training and testing, and a 10-fold cross validation
splits:

```{r}
# Create splits:
set.seed(123)
steph_split <- initial_split(base_steph_pbp, prop = 0.8, strata = scoring_play)
steph_train <- training(steph_split)
steph_test <- testing(steph_split)

steph_folds <- vfold_cv(steph_train, v = 10, strata = scoring_play)
```

The next step is to create a recipe for the model, we do this below,
with the game_id variable as the ID variable and one hot encoding for all
the nominal variables. The dependent variable is `scoring_play`, which will 
be a function of Curry's position on the court (coordinate_x_raw, coordinate_y_raw),
the opponent athlete, the quarter, the minutes and seconds left on the clock:

```{r}
steph_recipe <- recipe(scoring_play ~ ., data = steph_train) %>%
  update_role(game_id, new_role = "ID") %>% # Add ID step, for game level info
  step_novel(all_nominal(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes()) 

```

In addition to creating dummy variables for the nominal variables, we will also
account for the uniqueness of the play. That is, because there is an element 
of randomness in a basketball game, some of the players Curry's has been up 
against may not be in the test set or vice versa. To account for this, we will 
use `step_novel`. Now that we have the recipe, we can create the model workflow:

```{r}
#| echo: FALSE

# Model spec:
xgb_spec <- boost_tree(
  trees = 5000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry = tune(),
  learn_rate = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# Model workflow:
steph_workflow <- workflow() %>%
  add_recipe(steph_recipe) %>%
  add_model(xgb_spec)
```


```{r}
#| eval: false

# Model spec:
xgb_spec <- boost_tree(
  trees = 5000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry = tune(),
  learn_rate = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

# Model workflow:
steph_workflow <- workflow() %>%
  add_recipe(steph_recipe) %>%
  add_model(xgb_spec)

# Tuning grid:
xgb_grid <- grid_latin_hypercube(
  tree_depth(),
  min_n(),
  loss_reduction(),
  sample_size = sample_prop(),
  finalize(mtry(), steph_train),
  learn_rate(),
  size = 30
)

doParallel::registerDoParallel() # Activate parallel computing
set.seed(1234)

# Results of the tuning:
xgb_results <-  tune_grid(
  steph_workflow,
  resamples = steph_folds,
  grid = xgb_grid,
  control = control_grid(save_pred = TRUE)
)
```

Let's take a look at the fluctuation in terms of performance according 
to the model parameters:

```{r}
xgb_results %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  select(mean, mtry:sample_size) %>%
  pivot_longer(mtry:sample_size,
               values_to = "value",
               names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean)) +
  geom_point(alpha = 0.8, show.legend = FALSE, color="indianred") +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "AUC")+
  scale_color_brewer(palette = "Spectral")+
  theme_minimal()
```

It looks like the AUC metric is stabilising at around 0.75-0.80. With these
results in mind we finalise the workflow and perform the last fit:

```{r}

# Select best model and finalise workflow:

best_auc_model <- select_best(xgb_results,"roc_auc")

final_xgb <- finalize_workflow(
  steph_workflow,
  best_auc_model)

final_xgb %>%
  fit(data = steph_train) %>%
  extract_fit_parsnip() %>%
  vip(geom = "point") +
  theme_minimal()

final_res <- last_fit(final_xgb, steph_split)
```

The variable with the largest influence on whether Curry will score or not, is 
whether there is a direct opponent or not, i.e. there is a much higher likelihood
that the shot will go in if Curry is not against a specific opponent. The next two
most important variables are coordinate_x_raw and coordinate_y_raw, which denote
the position on the field. We can also plot the area under the curve and confusion
matrix:

```{r}
final_xgb %>%
  fit(data = steph_train) -> fit_steph

augment(fit_steph,steph_test)%>% 
  roc_curve(truth = scoring_play, .pred_TRUE) %>% 
  autoplot()
```

The model has a area under the curve of 0.78 and an accuracy of 0.73, the confusion
matrix is displayed below:

```{r}
collect_predictions(final_res) %>%
  conf_mat(scoring_play, .pred_class) %>%
  autoplot()+
  theme_minimal()
```

It looks like the model is better at predicting shots that don't go in, whereas
its accuracy is less clear cut when it comes to shots that actually went in. 

### Previously unseen data - 2024 NBA season

Up until now we have trained the model from the 2022 season up until the 2023 
season. Let's get the 2024 season data and make our predictions on that, to 
see how the model would perform with unseen data:

```{r}
# Get the data for the new season:
# So we filtered the data for the Golden State Warriors:

# NBA Team box data:
nba_team_box %>% 
  filter(season == 2024 & team_display_name == "Golden State Warriors") %>%
  select(game_id,opponent_team_name) -> game_ids
  
# NBA play-by-play data:  
nba_pbp %>%  
  filter(season == 2024) %>% 
  select(game_id,coordinate_x_raw,coordinate_y_raw,scoring_play,
         clock_minutes, athlete_id_2,clock_minutes,clock_seconds,qtr) %>% 
  left_join(player_ids,by=c("athlete_id_2"="athlete_id")) %>% 
  inner_join(game_ids, by=join_by(game_id)) %>% 
  select(-athlete_id_2) %>% 
  rename(oponent_athlete=athlete_display_name) %>% 
  mutate(oponent_athlete = if_else(is.na(oponent_athlete),"no direct oponent",oponent_athlete),
         scoring_play = fct_rev(as_factor(scoring_play))) -> base_steph_pbp_2024
```

Now we have Curry's 2024 season play-by-play data. Before predicting on the 2024
data, we fit the model to the whole training data:


```{r}

# Fit the model to the whole data: 

final_xgb %>%
    fit(data = base_steph_pbp) -> fit_steph

augment(fit_steph, base_steph_pbp_2024) %>%
  roc_auc(truth = scoring_play, .pred_TRUE)

# Assign to predictions to an object:

augment(fit_steph, base_steph_pbp_2024) -> steph_preds_2024
```

The model gives us an area under the curve of 80%. We can also break down the 
predictions according to specific sections of the basketball court and also 
overlay the time on the clock when the shot was made:

```{r}

# Plot the results:

steph_preds_2024 %>% 
  select(coordinate_x_raw,coordinate_y_raw,scoring_play,clock_minutes,.pred_class) %>%
  mutate(result=as.factor(if_else(.pred_class == TRUE,"made","missed")))->curry_shots_2024

subdata <- curry_shots_2024 %>% as.data.frame()
subdata$xx <- subdata$coordinate_x_raw-25
subdata$yy <- subdata$coordinate_y_raw-44

shotchart(data=subdata, x="xx", y="yy", z="clock_minutes", 
          num.sect=4, type="sectors", scatter=FALSE, result="result")+
  theme(legend.position = "bottom")+
  guides(fill = guide_legend(title = "Clock Minutes",title.position = "top",title.hjust = 0.5))
```

According to the model predictions, Curry is more likely to score when he is near
the basket. Similarly, the model predicts that somewhere between 40-50% of shots 
from the bottom left and right corners of the court will go in. 

Let's take a look at the model results, specifically focusing on false negatives and 
positives:

```{r}

# Plot the results:

steph_preds_2024 %>% 
  select(coordinate_x_raw,coordinate_y_raw,scoring_play,.pred_class,clock_minutes) %>%
  mutate(result=as.factor(if_else(scoring_play == TRUE,"made","missed"))) %>% 
  mutate(result2 = as.factor(case_when(
    scoring_play == TRUE & .pred_class == FALSE ~ "false negative",
    scoring_play == FALSE & .pred_class == TRUE ~ "false positive",.default = "Truth"))) %>% 
  filter(result2 != "Truth") %>% 
  select(-result) %>% 
  rename(result=result2) -> curry_shots_2024


subdata <- curry_shots_2024 %>% as.data.frame()
subdata$xx <- subdata$coordinate_x_raw-25
subdata$yy <- subdata$coordinate_y_raw-44

shotchart(data=subdata, x="xx", y="yy",z="result" ,type=NULL,scatter=TRUE)+
  theme(legend.position = "bottom")
```

It looks like the model predicts more false negatives for three pointers, i.e. 
along the three point line, as shown by the red points. In contrast, for false 
positives the concentration is mostly around the backboard, there is also a 
concentration of false negatives slightly to the right of the backboard, 
however for the most part it seems there are more false positives around that area. 
In terms of accuracy, the model achieves 75% accuracy, which is not great, but 
not terrible either.

## Conclusion & Future Research Directions

This project explores point guard trends in the NBA from season 2022 to 2024. It
begins by visualising and aggregatin 

















